Appendix: A fragile guardrail---The minimal effect of election fraud on protest intensity in non-democracies

### 2025-11-16 Output/authoritarian_only_pdf_pop_rev.Rmd
Final edits to variable descriptions

Using data from 659 non-democratic elections across 100 countries, from 1989 to 2012 (of which 156 experienced protest), this paper finds that governments that manipulate elections more heavily are likely to put down protests more quickly, especially when incumbents claim small vote-shares. This is inconsistent with the regime-weakness model of election manipulation, but consistent with a world in which election manipulation indicates sources of regime resilience. Given these results, it is likely that---while incumbents do not seek to provoke protest---the downside risk for incumbents that engage in election fraud in non-democracies has been overemphasized. This interpretation helps explain several empirical patterns at once: the relative paucity of electoral protests [@Brancati2016], the rarity of their success in overturning incumbent regimes, and governments' attempts to claim implausibly high margins of victory in conjunction with high levels of fraud. At a theoretical level, they also suggest that protest risk should be de-emphasized in research on the causes of election manipulation, in favor of alternative explanations which emphasize the role of front-line actors in delivering manipulation. Finally, they suggest a pessimistic outlook for democratization by election in authoritarian regimes.

@@

@@

It is likely that the theoretical mechanisms described above weaken as regimes become more liberal-democratic; in liberal democracies, patronage consolidation should decline and rule of law should increase by definition. These changes should make it more difficult for incumbents to manipulate elections, irrespective of protest risk. At the same time, the more liberal-democratic a state becomes, the more alternatives there are to protest as a means to check the executive. More independent courts and more active executive constraints have been shown to be associated with improved election integrity [@Harvey2022;@Bishop2016a]; it is where these factors are weakest---the electoral authoritarian regimes---where we should expect the deterrence logic to operate under a regime-weakness model. Finally, as a matter of measurement, protest duration is likely to be an increasingly imperfect measure of incumbent strength in more liberal-democratic settings, where protest is more commonplace and less immediately threatening to the regime. The case selection strategy here, excluding electoral and liberal democracies, is thus one that both fits the theory and makes for a stronger test of the regime-weakness model. Nevertheless, an expanded sample that includes electoral democracies provides similar results in the appendix. In total, there are 659 elections included in the dataset, across 100 countries. The date range is limited by the availability of data on protest from ECAV, which ranges from 1989 to 2012.

For models of protest-wave duration, as discussed in more detail below, I include only observations where a positive number of protests occurred; there are 211 such cases.

@@

ECAV records protests and other events related to an election, ranging from six months before to three months after the vote (or three months after the final round of a multi-round election). To qualify, an action must be contentious; rallies in support of the incumbent are excluded. I only include events labeled as protests, occupations, or blockades, excluding acts like shootings, bombings, and coups. If any post-election protests are recorded in the ECAV data for a particular election, the binary variable *protest initiation* is marked as 1. ECAV data is in event-day format; to capture *protest duration* I take the number of days between the last protest recorded in ECAV and the first, for each election.^[The duration of a protest wave with only one protest is taken to be one day.] For the survival models, protest movements that were ultimately defeated by the incumbent are considered 'dead' after the last day of protest. Protest movements that prevailed in unseating the incumbent are included, and are marked as 'alive.'

The large majority of elections experience no post-election protest: in `r nelda2.sub %>% filter(n.events.post == 0) %>% nrow()` of the `r nelda2.sub %>% nrow()` elections, no protests are recorded. Figure \@ref(fig:elex-hist) reports the distribution of election fraud across the dataset, as well as protest duration. As the figure shows, while election fraud is common among this set of regimes, protests are rare, and, when they do occur, are usually short in duration. These distributions---extensive election fraud combined with large numbers of non-protest---provides some early indication that the regime-weakness model has limited support. Still, approximately one-third of elections experienced some protest. Protest wave duration is modeled using Cox proportional hazards models, which have the following formula:

$$h(t | X)=h_0(t)* exp(\beta_fX_f+\beta_vX_v+ ... +\beta_pXp_)$$
The formula models the risk that a protest movement will fail at time $t$ given covariates $X$ is a function of the baseline hazard function, the degree of electoral fraud ($X_f$), the incumbent's vote-share ($X_v$) and a set of control variables.

Modeling protest duration rather than initiation helps reduce concerns about endogeneity. For instance, some regime-weakness models expect that incumbents will calibrate the level of election fraud to minimize risk of protest initiation. In that case, strong incumbents would manipulate more and weak incumbents less, but neither would systematically generate protest. Unpredictable and likely statistically insignificant correlations between fraud and protest initiation would be expected. Focusing on duration partially avoids this risk. Once protests begin (whether instrumental or multi-purpose), the die is cast. Authoritarian incumbents have strong incentives to defeat a public challenge to their legitimacy, while committed opponents have similar incentives to stand firm. Both should be expected to put significant effort into defeating / maintaining the protest movements, as a general rule, and the process by which the challenge is settled helps reveal previously hidden information about each side. Whether the incumbent manipulated 'too much' or 'too little' is now beside the point, and it is possible to observe whether incumbents who produce extensive election fraud are stronger or weaker than those that do not.

While strategic endogeneity is less of a concern for this research design, selection bias and confounding variables may still be. For instance, it could be the case that incumbents facing well-organized opposition-oriented civil society may manipulate less; at the same time, such well-organized opposition may be able to sustain protest movements longer. A similar logic might imply that incumbents should manipulate less when economic grievances are high (for fear of sparking protest), and that economic troubles might make protests more durable. Either case could produce the relationship expected in H2, but could be considered spurious. To reduce this risk, I include several potential controls for opposition strength, and for grievance. Controlling for these variables helps isolate the relationship between manipulation severity and protest duration, holding opposition strength and grievances constant.

As such, the results should be considered causally identified only to the extent that unknown or unobservable variables are not confounding the relationship between fraud and protest duration. While this problem is difficult to surmount with observational data, this research design offers three defenses. First, I control for known possible confounders, including the occurrence of a successful electoral protest in a nearby country (which might frighten incumbents into manipulating less and inspire opponents to hold the line longer). Second, I present two analyses in the appendix, which show that increased risk of protest leads incumbents to *intensify* their manipulation efforts, rather than diminish them. Thus a negative relationship between fraud and protest duration should indicate the direct hypothesized effect, rather than one that is driven by relatively weaker incumbents avoiding manipulation. Third, and relatedly, the research design pits two incompatible hypotheses against one another, rather than a single hypothesis against the null. A null relationship between fraud and protest duration might not be strong evidence against the regime-weakness model, but a negative relationship should be (after accounting for necessary controls).

```{r binarytab, echo=FALSE, message=FALSE, warning=FALSE}
count_table <- table(cut(nelda2.sub$winner.share.upd, breaks = c(0, .45, .55, 1), 
    include.lowest = TRUE, labels = c("Low", "Medium", "High")),
      cut(nelda2.sub$v2elirreg.inv, breaks =quantile(nelda2.sub$v2elirreg.inv, probs = c(0, 1/3, 2/3, 1), na.rm=T), 
    include.lowest = TRUE, labels = c("Low", "Medium", "High")),
      nelda2.sub$any.ecav.protest)

@@

As a plausibility check, I first examine patterns of protest initiation. The regime-weakness model expects initiation to increase with fraud, holding vote-share constant; while the regime-strength model expects a less consistent pattern. Table \@ref(tab:binarytab) categorizes elections into 3 x 3 bins based on incumbent vote-share and degree of electoral fraud, and reports the percentage of elections in each category that were followed by protest.  For this table, vote-share is considered low if it less than 45%, medium between 45% and 55%, and high if greater than 55%. The middle category is intended to capture a 'close' election. Election fraud is divided into terciles. These categories can be thought of as relatively clean (*election fraud* is between -2.33 and 0.5), moderately manipulated (0.5 to 1.46) and heavily manipulated (1.46 to 3.2). Two patterns appear in the data. First, protest initiation is most likely when incumbent vote-share is low and fraud is high. This is expected under both models. For the regime-weakness model, this is taken as a sign that opposition actors launch protests instrumentally against the weakest (i.e. least popular) incumbents. For the regime-strength model, this pattern could emerge as electoral manipulation focuses mass grievances. However, we do not observe a consistent increase in protest likelihood as manipulation increases, as the regime-weakness model would predict. Instead, we see relatively consistent rates of protest initiation at medium vote-share across levels of manipulation. And, at high incumbent vote-shares, there is even a decrease in protest initiation as manipulation increases from moderate to high. This comports with the mixed evidence found in prior research on protest initiation and manipulation, and is evidence that the regime-strength model is on a plausible foundation (specifically, that protests can occur even when the incumbent is predicted to be 'strong' under either model). Consider the cell where vote-share is high and election fraud is low---this is the strongest (i.e. clearly popular) type of regime in the regime-weakness model. Nevertheless, more than a quarter of such cases are protested. This is not so different from the cell which is most likely to contain strong regimes in the regime-strength model (high vote-share, high manipulation), in which thirty-two percent of cases launch protest. While not dispositive, this result indicates that the regime-strength model and regime-weakness models are equally compatible with patterns of protest initiation, as expected.

@@

I operationalize electoral manipulation using this variable, rather than alternative measures of election integrity, because fraud of this kind is thought to be most likely to trigger protest [@Linebarger2020]---it is often incumbents' tool of last resort [@Sjoberg2016], may not mobilize real voters [@Harvey2020], and can be more damaging to legitimacy than pre-election manipulation [@birch2011;@Szakonyi2021]. This measure of election integrity should thus serve as a strong test of the protest-oriented model. It also matches the conception of election manipulation used in much of the relevant formal literature.^[For example, election manipulation is treated as post-hoc adjustment of the results in Little et al [-@Little2015], Little [-@Little2012], Magaloni [-@magaloni2010], and Fearon [-@fearon_self-enforcing_2011], among others.] An alternative measure is used as a robustness check in the appendix.

@@

First, I include the *number of pre-election protests* counted in the ECAV dataset, which is likely to be an important possible confounder; more numerous pre-election protest may affect the level of election fraud (as the deterrence logic predicts) as well as being directly tied to the likelihood of post-election protest. Next, the models of protest duration include a measure of the *median size of protest*, calculated using ECAV data.^[ECAV data codes protest size as a categorical variable, ranging from 1 (< 10 participants) to 5 (> 10,000). I take the median size category of all the events in the protest wave as a measure of average protest size.] This is important as a control for opposition strength, as well as for the possibility that incumbents might tolerate smaller protests. Since non-democratic governments may be able to adjust the electoral calendar strategically, to disrupt opposition coordination [@nygardTimingMattersImpact2020], I include a variable from NELDA that indicates *off-schedule elections*.

Since higher repressiveness may weaken or intensify opposition [@chenowethStateRepressionNonviolent2017], I include a lagged measure of physical integrity index from V-Dem (v2x_clphy), and its square term. Indicators of government *control over civil society* (the 'v2cseeorgs' variable from V-Dem), the proportion of national political parties with *permanent party organizations* (V-Dem's 'v2psorgs'), *legislative constraints* on the executive (V-Dem variable 'v2xlg_legcon', and the availability of *alternative sources of information* ('v2x_freexp_altinf' from V-Dem) are also employed as straightforward indicators of opposition strength. Since urban areas can make electoral manipulation more difficult [@ziblatt2009;@larreguy2016parties] while also making protest more likely [@Wallace2013a], I also include data on *urbanization* taken from the United Nations [@DepartmentofEconomicandSocialAffairsPopulationDivision2019].

@@

In addition, the literature on the diffusion of electoral protest holds that successful electoral protest in one country increases the likelihood of electoral protest in similar countries nearby due to opposition learning [@Bunce2010;@Beissinger_2007]. Such a dynamic could plausibly increase the likelihood and severity of protest in a nearby country, while also reducing the incumbent's use of fraud. I control for this using each country's geopolitical region as coded in Teorell et al [-@qog_teorell], in conjunction with binary variables indicating a successful protest movement from NELDA. A 'successful protest' is recorded if a protest movement resulted in either a new election being run or the incumbent stepping down. The *regional diffusion* dummy variable takes on a value of one for all countries in the same region as a successful protest, for elections within one calendar year of the successfully overturned election.^[For example, the variable takes on a positive value for elections held in the post-communist region, within one year of the September 24, 2000 election in Serbia that sparked the 'Bulldozer Revolution'. All told there are 14 successful electoral protests in the dataset: Bulgaria 1990, Burkina Faso 1997, Lesotho 1998, Cote d'Ivoire 2000 and 2010, Peru 2000, Serbia 2000, Mali 2002, Georgia 2003, Ukraine 2004, Azerbaijan 2005, Kyrgyzstan 2005, Haiti 2006, and Cameroon 2007.]

@@

@@

cox1 <- coxph(Surv(d.events.num, surv_status) ~  legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag +  winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct  +  unscheduled_election + regional_success_lag + v2elirreg.inv   + n.events.post , data = nelda2.sub %>% filter(inc.lose != "1" &  n.events.post > 0))
summary(cox1)

cox1.cen <- coxph(Surv(d.events.num, surv_status_censor_assume) ~  legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag +  winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct  +  unscheduled_election + regional_success_lag + v2elirreg.inv   + n.events.post , data = nelda2.sub %>% filter(inc.lose == "0" & n.events.post > 0))
summary(cox1.cen)

@@

cox2 <- coxph(Surv(d.events.num, surv_status) ~  legcon.1lag + altinf.1lag  +  osorg.1lag + gdpgro.1lag +  + v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + unscheduled_election + regional_success_lag + n.events.post, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0))
summary(cox2)
p.duration.margin <- sjPlot::plot_model(cox2, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [-.29, 2.2]")) #This is the 10th and 90th quantiles
p.duration.margin <- p.duration.margin +  labs(x = "Incumbent vote-share", y = "Hazard ratio", title = "Hazard ratio for protest movement end", color = "Election fraud") + theme_bw()

cox2.cen <- coxph(Surv(d.events.num, surv_status_censor_assume) ~  legcon.1lag + altinf.1lag  +  osorg.1lag + gdpgro.1lag + v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + unscheduled_election + regional_success_lag + n.events.post, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0))
summary(cox2.cen)

cox3.cen <- coxph(Surv(d.events.num, surv_status_censor_assume) ~  legcon.1lag + altinf.1lag  +  osorg.1lag + gdpgro.1lag +  v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + unscheduled_election + regional_success_lag + n.events.post + median.participants, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0))
summary(cox3.cen)

>>>>

Using data from 659 non-democratic elections across 100 countries, from 1989 to 2012 (of which 211 experienced protest), this paper finds that governments that manipulate elections more heavily are likely to put down protests more quickly, especially when incumbents claim small vote-shares. This is inconsistent with the regime-weakness model of election manipulation, but consistent with a world in which election manipulation indicates sources of regime resilience. Given these results, it is likely that---while incumbents do not seek to provoke protest---the downside risk for incumbents that engage in election fraud in non-democracies has been overemphasized. This interpretation helps explain several empirical patterns at once: the relative paucity of electoral protests [@Brancati2016], the rarity of their success in overturning incumbent regimes, and governments' attempts to claim implausibly high margins of victory in conjunction with high levels of fraud. At a theoretical level, they also suggest that protest risk should be de-emphasized in research on the causes of election manipulation, in favor of alternative explanations which emphasize the role of front-line actors in delivering manipulation. Finally, they suggest a pessimistic outlook for democratization by election in authoritarian regimes.

@@

It is likely that the theoretical mechanisms described above weaken as regimes become more liberal-democratic; in liberal democracies, patronage consolidation should decline and rule of law should increase by definition. These changes should make it more difficult for incumbents to manipulate elections, irrespective of protest risk. At the same time, the more liberal-democratic a state becomes, the more alternatives there are to protest as a means to check the executive. More independent courts and more active executive constraints have been shown to be associated with improved election integrity [@Harvey2022;@Bishop2016a]; it is where these factors are weakest---the electoral authoritarian regimes---where we should expect the deterrence logic to operate under a regime-weakness model. Finally, as a matter of measurement, protest duration is likely to be an increasingly imperfect measure of incumbent strength in more liberal-democratic settings, where protest is more commonplace and less immediately threatening to the regime.

@@

The case selection strategy here, excluding electoral and liberal democracies, is one that both fits the theory and makes for a stronger test of the regime-weakness model. Nevertheless, an expanded sample that includes a set of electoral democracies provides similar results in the appendix. In total, there are 659 elections included in the dataset, across 100 countries, with 211 experiencing post-election protest. The date range is limited by the availability of data on protest from ECAV, which ranges from 1989 to 2012.

@@

ECAV records protests and other events related to an election, ranging from six months before to three months after the vote (or three months after the final round of a multi-round election). To qualify, an action must be contentious; rallies in support of the incumbent are excluded. I only include events labeled as protests, occupations, or blockades, excluding acts like shootings, bombings, and coups. ECAV data is in event-day format; to capture *protest duration* I take the number of days between the last protest recorded in ECAV and the first, for each election.^[The duration of a protest wave with only one protest is taken to be one day.] Protest wave duration is modeled using Cox proportional hazards models, which have the following formula:

$$h(t | X)=h_0(t)* exp(\beta_fX_f+\beta_vX_v+ ... +\beta_pXp_)$$
The formula models the risk that a protest movement will fail at time $t$ given covariates $X$ is a function of the baseline hazard function, the degree of electoral fraud ($X_f$), the incumbent's vote-share ($X_v$) and a set of control variables. For the survival models, protest movements that were ultimately defeated by the incumbent are considered 'dead' after the last day of protest. Protest movements that prevailed in unseating the incumbent are included, and are marked as 'alive.'

Modeling protest duration rather than initiation helps reduce concerns about endogeneity. For instance, some regime-weakness models expect that incumbents will calibrate the level of election fraud to minimize risk of protest initiation. In that case, strong incumbents would manipulate more and weak incumbents less, but neither would systematically generate protest. Unpredictable and likely statistically insignificant correlations between fraud and protest initiation would be expected. Focusing on duration partially avoids this risk. Once protests begin (whether instrumental or multi-purpose), authoritarian incumbents have strong incentives to defeat a public challenge to their legitimacy, while committed opponents have similar incentives to stand firm. Both should be expected to put significant effort into defeating / maintaining the protest movements, as a general rule, and the process by which the challenge is settled helps reveal previously hidden information about each side; it is possible to observe whether incumbents who produce extensive election fraud are stronger or weaker than those that do not.

While strategic endogeneity is less of a concern for this research design, selection bias and confounding variables may still be. For instance, it could be the case that incumbents who face well-organized opposition-oriented civil society may manipulate less; at the same time, such well-organized opposition may be able to sustain protest movements longer. A similar logic might imply that incumbents should manipulate less when economic grievances are high (for fear of sparking protest), and that economic troubles might make protests more durable. Either case could produce the relationship expected in H2, but could be considered spurious. I take several steps to reduce this risk.

First, I control for measures of opposition strength and grievances, to help isolate the relationship between manipulation severity and protest duration. Still, it is possible that unknown or unobservable variables may confound the relationship between fraud and protest duration. While this problem is difficult to surmount with observational data, this research design offers three defenses. Next, I present two analyses in the appendix, which show that increased risk of protest leads incumbents to *intensify* their manipulation efforts, rather than diminish them. Thus a negative relationship between fraud and protest duration should indicate the direct hypothesized effect, rather than one that is driven by relatively weaker incumbents selecting out of manipulation. Third, and relatedly, the research design pits two incompatible hypotheses against one another, rather than a single hypothesis against the null. A null relationship between fraud and protest duration might not be strong evidence against the regime-weakness model, but a negative relationship should be (after accounting for necessary controls).

```{r binarytab, echo=FALSE, message=FALSE, warning=FALSE, eval=FALSE}
count_table <- table(cut(nelda2.sub$winner.share.upd, breaks = c(0, .45, .55, 1), 
    include.lowest = TRUE, labels = c("Low", "Medium", "High")),
      cut(nelda2.sub$v2elirreg.inv, breaks =quantile(nelda2.sub$v2elirreg.inv, probs = c(0, 1/3, 2/3, 1), na.rm=T), 
    include.lowest = TRUE, labels = c("Low", "Medium", "High")),
      nelda2.sub$any.ecav.protest)

@@

<!-- As a plausibility check, I first examine patterns of protest initiation. The regime-weakness model expects initiation to increase with fraud, holding vote-share constant; while the regime-strength model expects a less consistent pattern. Table \@ref(tab:binarytab) categorizes elections into 3 x 3 bins based on incumbent vote-share and degree of electoral fraud, and reports the percentage of elections in each category that were followed by protest.  For this table, vote-share is considered low if it less than 45%, medium between 45% and 55%, and high if greater than 55%. The middle category is intended to capture a 'close' election. Election fraud is divided into terciles. These categories can be thought of as relatively clean (*election fraud* is between -2.33 and 0.5), moderately manipulated (0.5 to 1.46) and heavily manipulated (1.46 to 3.2). Two patterns appear in the data. First, protest initiation is most likely when incumbent vote-share is low and fraud is high. This is expected under both models. For the regime-weakness model, this is taken as a sign that opposition actors launch protests instrumentally against the weakest (i.e. least popular) incumbents. For the regime-strength model, this pattern could emerge as electoral manipulation focuses mass grievances. However, we do not observe a consistent increase in protest likelihood as manipulation increases, as the regime-weakness model would predict. Instead, we see relatively consistent rates of protest initiation at medium vote-share across levels of manipulation. And, at high incumbent vote-shares, there is even a decrease in protest initiation as manipulation increases from moderate to high. This comports with the mixed evidence found in prior research on protest initiation and manipulation, and is evidence that the regime-strength model is on a plausible foundation (specifically, that protests can occur even when the incumbent is predicted to be 'strong' under either model). Consider the cell where vote-share is high and election fraud is low---this is the strongest (i.e. clearly popular) type of regime in the regime-weakness model. Nevertheless, more than a quarter of such cases are protested. This is not so different from the cell which is most likely to contain strong regimes in the regime-strength model (high vote-share, high manipulation), in which thirty-two percent of cases launch protest. While not dispositive, this result indicates that the regime-strength model and regime-weakness models are equally compatible with patterns of protest initiation, as expected.

-->

@@

I operationalize electoral manipulation using this variable, rather than alternative measures of election integrity, because fraud of this kind is thought to be most likely to trigger protest [@Linebarger2020]---it is often incumbents' tool of last resort [@Sjoberg2016], may not mobilize real voters [@Harvey2020], and can be more damaging to legitimacy than pre-election manipulation [@birch2011;@Szakonyi2021]. This measure of election integrity should thus serve as a strong test of the protest-oriented model. It also matches the conception of election manipulation used in much of the relevant formal literature.^[For example, election manipulation is treated as post-hoc adjustment of the results in Little et al [-@Little2015], Little [-@Little2012], Magaloni [-@magaloni2010], and Fearon [-@fearon_self-enforcing_2011], among others.] Two alternative measures for manipulation are used as robustness checks in the appendix.

@@

First, I include the *number of pre-election protests* counted in the ECAV dataset, which is likely to be an important possible confounder; more numerous pre-election protest may affect the level of election fraud (as the deterrence logic predicts) as well as being directly tied to the likelihood of post-election protests. Next, I control for the *median size of protest*, calculated using ECAV data.^[ECAV data codes protest size as a categorical variable, ranging from 1 (< 10 participants) to 5 (> 10,000). I take the median size category of all the events in the protest wave as a measure of average protest size.] This is important as a control for opposition strength, as well as for the possibility that incumbents might tolerate smaller protests. Since non-democratic governments may be able to adjust the electoral calendar strategically, to disrupt opposition coordination [@nygardTimingMattersImpact2020], I include a variable from NELDA that indicates *off-schedule elections*.

Since higher repressiveness may weaken *or* intensify opposition [@chenowethStateRepressionNonviolent2017], I include a lagged measure of *physical integrity* index from V-Dem (v2x_clphy), and its square term. Indicators of government *control over civil society* (the 'v2cseeorgs' variable from V-Dem), the proportion of national political parties with *permanent party organizations* (V-Dem's 'v2psorgs'), *legislative constraints* on the executive (V-Dem variable 'v2xlg_legcon', and the availability of *alternative sources of information* ('v2x_freexp_altinf' from V-Dem) are also employed as straightforward indicators of opposition strength. Since urban areas can make electoral manipulation more difficult [@ziblatt2009;@larreguy2016parties] while also making protest more likely [@Wallace2013a], I also include data on *urbanization* taken from the United Nations [@DepartmentofEconomicandSocialAffairsPopulationDivision2019].

@@

In addition, the literature on the diffusion of electoral protest holds that successful electoral protest in one country increases the likelihood of electoral protest in similar countries nearby due to opposition learning [@Bunce2010;@Beissinger_2007]. Such a dynamic could plausibly increase the severity of protest in a nearby country, while also reducing the incumbent's use of fraud. I control for this using each country's geopolitical region as coded in Teorell et al [-@qog_teorell], in conjunction with binary variables indicating a successful protest movement from NELDA. A 'successful protest' is recorded if a protest movement resulted in either a new election being run or the incumbent stepping down. The *regional diffusion* dummy variable takes on a value of one for all countries in the same region as a successful protest, for elections within one calendar year of the successfully overturned election.^[For example, the variable takes on a positive value for elections held in the post-communist region, within one year of the September 24, 2000 election in Serbia that sparked the 'Bulldozer Revolution'. All told there are 14 successful electoral protests in the dataset: Bulgaria 1990, Burkina Faso 1997, Lesotho 1998, Cote d'Ivoire 2000 and 2010, Peru 2000, Serbia 2000, Mali 2002, Georgia 2003, Ukraine 2004, Azerbaijan 2005, Kyrgyzstan 2005, Haiti 2006, and Cameroon 2007.]

@@

The large majority of elections experience no post-election protest: in `r nelda2.sub %>% filter(n.events.post == 0) %>% nrow()` of the `r nelda2.sub %>% nrow()` elections, no protests are recorded. Figure \@ref(fig:elex-hist) reports the distribution of election fraud across the dataset, as well as protest duration. As the figure shows, while election fraud is common among this set of regimes, protests are rare, and, when they do occur, are usually short in duration. These distributions---extensive election fraud combined with large numbers of non-protest---provides some early indication that the regime-weakness model has limited support. Still, approximately one-third of elections experienced some protest.

@@

cox1 <- coxph(Surv(d.events.num, surv_status) ~  lji.lag + legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag +  winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct  +  unscheduled_election + regional_success_lag + v2elirreg.inv   + n.events.post , data = nelda2.sub %>% filter(inc.lose != "1" &  n.events.post > 0))
summary(cox1)

cox1.cen <- coxph(Surv(d.events.num, surv_status_censor_assume) ~ lji.lag + legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag +  winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct  +  unscheduled_election + regional_success_lag + v2elirreg.inv   + n.events.post , data = nelda2.sub %>% filter(inc.lose == "0" & n.events.post > 0))
summary(cox1.cen)

@@

cox2 <- coxph(Surv(d.events.num, surv_status) ~ lji.lag + legcon.1lag + altinf.1lag  +  osorg.1lag + gdpgro.1lag +  + v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + unscheduled_election + regional_success_lag + n.events.post, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0))
summary(cox2)
p.duration.margin <- sjPlot::plot_model(cox2, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [-.29, 2.2]")) #This is the 10th and 90th quantiles
p.duration.margin <- p.duration.margin +  labs(x = "Incumbent vote-share", y = "Hazard ratio", title = "Hazard ratio for protest movement end", color = "Election fraud") + theme_bw()

cox2.cen <- coxph(Surv(d.events.num, surv_status_censor_assume) ~ lji.lag +  legcon.1lag + altinf.1lag  +  osorg.1lag + gdpgro.1lag + v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + unscheduled_election + regional_success_lag + n.events.post, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0))
summary(cox2.cen)

cox3.cen <- coxph(Surv(d.events.num, surv_status_censor_assume) ~ lji.lag +  legcon.1lag + altinf.1lag  +  osorg.1lag + gdpgro.1lag +  v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + unscheduled_election + regional_success_lag + n.events.post + median.participants, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0))
summary(cox3.cen)
